---
title: "Available methods"
author: "Indrajeet Patil"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig.align: 'center'
    fig.asp: 0.618
    dpi: 300
    toc: true
    warning: FALSE
    message: FALSE
vignette: >
  %\VignetteIndexEntry{Available methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The goal of `broomExtra` is to provide helper functions that assist in data
analysis workflows involving regression analyses. The goal is to combine the
functionality offered by different set of packages through a common syntax to
return tidy tibbles containing model parameters and summaries.

The package internally relies on the following packages that I contribute to:

  - `broom`
  - `broom.mixed`
  - `parameters`
  - `performance`

Here is a summary of supported methods across these ecosystems.

# The `broom` ecosystem

Currently, `S3` methods for mixed-effects model objects are included in the
`broom.mixed` package, while the rest of the object classes are included in the
`broom` package. This means that you constantly need to keep track of the class
of the object (e.g., "if it is `merMod` object, use
`broom.mixed::tidy()`/`broom.mixed::glance()`/`broom.mixed::augment()`, but if
it is `polr` object, use `broom::tidy()`/`broom::glance()`/`broom::augment()`").
Using generics from `broomExtra` means you no longer have to worry about this,
as calling `broomExtra::tidy()`/`broomExtra::glance()`/`broomExtra::augment()`
will search the appropriate method from these two packages and return the
results.

Here are all available methods currently available in `broom` ecosystem:

```{r list, echo = FALSE, message = FALSE}
# loading the needed libraries
library(dplyr)
library(broom)
library(broom.mixed)

# function to extract methods
method_df <- function(method_name) {
  m <- as.vector(methods(method_name))
  dplyr::tibble(
    class = gsub(
      x = m,
      pattern = paste(method_name, "[.]", sep = ""),
      replacement = ""
    ),
    !!method_name := "x"
  )
}

# preparing a table
method_df("tidy") %>%
  dplyr::left_join(x = ., y = method_df("glance")) %>%
  dplyr::left_join(x = ., y = method_df("augment")) %>%
  dplyr::mutate_all(.tbl = ., .funs = tidyr::replace_na, "") %>%
  knitr::kable()
```

# The `easystats` ecosystem

Here are all available methods currently available in `easystats` ecosystem:

```{r}
insight::supported_models()
```